---
title: "Exploring Fuel Efficiency with mtcars"
author: "Rshan Hluf"
date: today
format: html
---
## Exploring the mtcars Dataset

In this post, I explore the relationship between fuel efficiency, vehicle weight, horsepower, and number of cylinders using R.

The data come from the [`mtcars` dataset](https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/mtcars.html), which is included with R and contains information about 32 automobiles from the 1974 *Motor Trend* US magazine.
## Preparing the Data

```{r}
#| message: false
#| warning: false

cars <- mtcars
cars$cyl <- factor(cars$cyl)

head(cars)
```
## Fuel Efficiency by Number of Cylinders

Next, I compare the average miles per gallon for cars with 4, 6, and 8 cylinders.

```{r}
mean_mpg <- aggregate(mpg ~ cyl, data = cars, FUN = mean)
mean_mpg
```
## Fuel Efficiency and Vehicle Weight

The following figure shows the relationship between vehicle weight and fuel efficiency.

```{r}
#| label: fig-mpg-weight
#| fig-cap: "Relationship between vehicle weight and fuel efficiency in the mtcars dataset."

plot(
  cars$wt,
  cars$mpg,
  xlab = "Weight (1000 lbs)",
  ylab = "Miles per gallon",
  main = "Fuel Efficiency vs. Vehicle Weight"
)

abline(lm(mpg ~ wt, data = cars), lwd = 2)
```
## Modeling Fuel Efficiency

Finally, I fit a linear regression model to examine how vehicle weight and horsepower are associated with fuel efficiency.

```{r}
model <- lm(mpg ~ wt + hp, data = cars)
summary(model)
```
## Findings

The results suggest that heavier cars tend to have lower fuel efficiency. In the regression model, vehicle weight has an estimated coefficient of -3.88, meaning that higher weight is associated with lower miles per gallon when horsepower is held constant. Horsepower also has a negative estimated association with fuel efficiency.

The model has an R-squared value of approximately 0.83, indicating that vehicle weight and horsepower together explain about 83% of the variation in miles per gallon among the cars in this dataset. These results describe associations in the `mtcars` data and should not be interpreted as evidence of causation.